Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-10 fib2.tex
more edits

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ observables arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the oservables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{T\in \Lambda}\proj{\vac}_T\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_T$ the projector to charge $\vac$ at tile $T$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.
%We identify this ground space as the codespace of our model. 
Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, we use heuristics
to minimize non-local braid moves, which helps to
maintain the state as a separable product.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
%However, for finite lattice size, this regime is somewhere
%beyond the decoder threshold, which is the 
%point at which the decoder creates nontrivial (extensive) loops.
%\stf{CGB and I are still not confident in the previous sentence or the next one}
%We can view the system state as a tensor network, on which
%the decoder is performing an (efficiently simulable) contraction.
%This enables simulation of error correction in
%the area around decoder threshold for lattice sizes up to $L=128.$
%\dude{more polish needed...}

>>>>

% We need to relate this to discrete models: toric code, string nets, etc.

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
%We tile the manifols with  a finite set of $L\times L$ Wilson loops arranged in
These \emph{tiles} will be the observables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{T\in \Lambda}\proj{\vac}_T\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_T$ the projector to charge $\vac$ at tile $T$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.
%We identify this ground space as the codespace of our model.

Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

In the regime where the action of noise plus decoder does not percolate it is
clear that the simulation is efficient.
However at some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics %for finite lattice size, we use heuristics
that minimize non-local braid moves, which helps to
maintain the state as a separable product beyond the
sub-percolated regime.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
%However, for finite lattice size, this regime is somewhere
%beyond the decoder threshold, which is the 
%point at which the decoder creates nontrivial (extensive) loops.
%\stf{CGB and I are still not confident in the previous sentence or the next one}
%We can view the system state as a tensor network, on which
%the decoder is performing an (efficiently simulable) contraction.
%This enables simulation of error correction in
%the area around decoder threshold for lattice sizes up to $L=128.$
%\dude{more polish needed...}
